---  
title: "SCAPES - Data Management Plan"  
author: "SCAPES study team"  
date: "2024-02-05"  # Replace with the current date  
output:  
  pdf_document:  
    toc: true
    toc_depth: 4
    number_sections: true
    latex_engine: pdflatex
    includes:  
      in_header: styles.tex  
---  

# Project data 

```{r child = "project_data.Rmd"}
```

# Planned project outputs

```{r child = "project_outputs.Rmd"}
```

# Qualitative data  

## Participatory Rural Assessment  

How will data be collected?    
What form will this data take?  
How/where will it be stored?  
What will the data include?  
What are the variables/parameters of the data?  
How will data be cleaned/processed?  
How will the data be anonymised/participants protected? Will data be shared? If so, how?  

# Quantitative data  

## Community questionnaires  

Questionnaires have been designed as XLSForms and are implemented using the Kobo platform. Data is collected on local devices and sent to the KoboToolbox server (<https://kf.kobotoolbox.org/>) for the SCAPES project. Access can be provided for data entry, modification and download. An unprocessed data set can be downloaded directly from the KoboToolbox server.  

Alternatively, a project has been created to pull the data and format it using R.  

### R Repository for data processing  

The `robotoolbox` R package is used to download the data locally through the Kobo API. This approach uses the `dm` R package to handle the grouped repeats in the questionnaire and the `haven` R package to handle variable structure (e.g., labelled factors).  

Within the `dm` structure repeats within the questionnaire form unique rows in a linked dataframe.  

``` {r dm-description, message = FALSE, warning = FALSE}  

library(here)  
library(dm)  
library(dplyr)  
library(readr)  

dm <- read_rds(here("household_questionnaire", "data", "h_data", "h_data_dm_2024-02-05.rds"))  

knitr::kable(tibble(`Table names` = names(dm),  
                    `Number of records` = sapply(dm, function(x) nrow(x)),  
                    `Links to households` = sapply(dm, function(x)   
                      length(unique(x$`_parent_index`)))))  

dm_draw(dm)  

```  

Within questionnaires these repeats may be conditional and so each main questionnaire may be associated with none or many indexed repeats.  

### Household questionnaire:  

```{r child = "hh_questionnaires_section.Rmd"}
```

### Household questionnaire data checks:

```{r child = "hh_questionnaire_checks.Rmd"}
```

### Individual questionnaire:  

```{r child = "individual_questionnaires_section.Rmd"}
```


### Individual questionnaire data checks:

```{r child = "individual_questionnaires_checks.Rmd"}
```

Linking IDs to samples?  

### Human serology  

### Rodent trapping  

### Rodent serology/PCR  

### Other  
